Parallel Multipoint Approximation Method for Large Scale Optimization Problems \thanks{This study was supported by the Russian Science Foundation, project No. 16-11-10150.}

### 2017-11-22 2018/PAVT_Toropov/PAVT2018_MAM.tex
Table 1

\usepackage[T2A]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[russian,english]{babel}

@@

\Russian
In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered. In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}. This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

@@

However, when the number of variables becomes of the order of $10^3$, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function 
values remains constant). Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 2000 times (from 4 up to 9000 seconds) when increasing the number of variables from 100 up to  1000. Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the 
present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.

@@

\begin{table}
	\caption{Time and speedup}
	\label{tab:1}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & MAM time & PMAM time & Speedup \\
		\hline\noalign{\smallskip}
		100 & 48 &  34 &   \\
		200 & 48 &  34 &   \\
		500 & 48 &  34 &   \\
		1000 & 107479  &  19469 &   \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Function evaluations}
	\label{tab:2}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
		\hline\noalign{\smallskip}
		100 & 10 &  2201 & 9901 \\
		200 & 10 &  4402 & 28143 \\
		500 &  &   &  \\
		1000 &  &   &  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

>>>>

%\usepackage[T2A]{fontenc}
%\usepackage[utf8x]{inputenc}
%\usepackage[russian,english]{babel}

@@

In the present paper, the multipoint approximation method (MAM \cite{Toropov1989,Toropov1992,ToropovFilatov1993}) and its application to the large scale optimization problems are considered. In problems with a large (in the order of hundreds) number of design variables MAM proved to be efficient, e.g. in turbomachinery applications \cite{ShahparPolynkinToropov2008,PolynkinToropovShahpar2008,PolynkinToropovShahpar2010}. This method is an iterative optimization technique based on mid-range approximations built in trust regions. A trust region is a sub-domain of the design space in which a set of design points, produced according to a small-scale design of experiments (DoE), are evaluated. These and a subset of previously evaluated design points are used to build metamodels of the objective and constraint functions that are considered to be valid within a current trust region. The trust region will then translate and change size as optimization progresses. The trust region strategy has gone through several stages of development to account for the presence of numerical noise in the response function values \cite{KeulenToropovMarkine1996,ToropovKeulenMarkine1996} and occasional simulation failures \cite{ToropovMarkineHolden1999}. The mid-range approximations used in the trust regions, as originally suggested in \cite{Toropov1989} for structural optimization problems, are intrinsically linear functions (i.e. nonlinear functions that can be led to a linear form by a simple transformation) for individual sub-structures, and assembly of them for the whole structure. This was enhanced by the use of gradient-assisted metamodels \cite{ToropovFilatov1993}, use of simplified numerical models that is also termed a multi-fidelity approach,\cite{ToropovMarkine1996} and the use of analytical models derived by Genetic Programming \cite{ToropovAlvarez1998}. One of the recent developments \cite{PolynkinToropov2012} involved the use of approximation assemblies, i.e. a two stage approximation building process that is conceptually similar to the original one used in \cite{Toropov1989} but is free from the limitation that lower level approximations are linked to individual substructures.

@@

However, when the number of variables becomes of the order of $10^3$, the work of the sequential part of MAM begins to affect the total problem solving time essentially (assuming the time of computing the objective function 
values remains constant). Thus, for the problem considered in Section 4 the time of execution of a single iteration of the method increased more than 4000 times (from 1.5 up to 6000 seconds) when increasing the number of variables from 100 up to  1000. Therefore, another approach to the parallelization of the algorithm has been applied within the framework of the 
present study: namely, the computational rules of MAM  providing the construction of the approximating, the solving of the approximated problem, and the choice of the next trust region (Steps 2b, 2c, and 2d of the algorithm) have been parallelized.

@@

\begin{table}
	\caption{Time and speedup}
	\label{tab:1}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $T_{MAM}$ & $T_{PMAM}$ & Speedup \\
		\hline\noalign{\smallskip}
		100 & 15.4 & 11.9 & 1.3  \\
		200 & 167 &  89 &  1.9 \\
		500 & 5151 &  1520 & 3.4  \\
		1000 & 67674 &  13771 & 4.9  \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}

\begin{table}
	\caption{Function evaluations}
	\label{tab:2}
	\center
	\begin{tabular}{cccc}
		\hline\noalign{\smallskip}
		$N$ & $I_{MAM}$  & $F_{MAM}$ & $F_{SQP}$ \\
		\hline\noalign{\smallskip}
		100 & 10 &  2201 & 9901 \\
		200 & 10 &  4402 & 28143 \\
		500 &  &   &  115734 \\
		1000 &  &   & 305308 \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table}
